---
title: "jlong"
output: github_document
---

The `R` package **jlong** provides a main function `jlong` to connect two 
    linear mixed-effects models via a joint model method.
    
## Installation

You can install the development version at [GitHub][github-url] using
the following code,

```R
if (!require(devtools)) install.packages("devtools")
devtools::install_github("https://github.com/ytalau/jlong")
```

```{r}
library(jlong)
```


## Simulated Data

### Submodel/Biomarker Data

There are some baseline variables contained in the submodel data:

- `biomk`: biomarker value
- `id`: unique subject identifier
- `time`: the visit time
- `gender`: 0 corresponds to female and 1 corresponds to male.
- `baseline_age`: age at time = 0



```{r}
head(example_dat$biomk)
```

There are additional baseline variables contained in the primary model data:

- `score`: clinical outcome score
- `id`: unique subject identifier which needs to match the `id` in the submodel
- `time`: the visit time which could be different from the visit time in submodel
- `gender`: 0 corresponds to female and 1 corresponds to male.
- `baseline_age`: age at time = 0
- `baseline_edu` number of years of education completed at time = 0
- `gene`: a binary variable indicating the existence of certain genes

```{r}
head(example_dat$primary)
```



## True Model: Primary model has random slope and random intercept

The `jlong` function is used to jointly fit two linear mixed models. 
There are five argument needed for the `jlong` to function.

- `data.primod`: a data frame / data.table for the primary model
- `data.submod`: a data frame / data.table for the biomarker model
- `formula.primod`: an object of class "formula" (or one that can be coerced to that class): a symbolic description of the primary model to be fitted; 
    the syntax of the formula is similar to the functions in `lme4` package
- `formula.submod`: an object of class "formula" (or one that can be coerced to that class): a symbolic description of the sub model to be fitted 
    similar to the functions in `lme4` package
- `var.shared`: a string specifying the variable name shared in both models

```{r}
res2 <- jlong(data.primod = example_dat$primary,
                  data.submod = example_dat$biomk,
                  formula.primod = score ~ gender + baseline_age + 
                      baseline_edu + gene + time  + gene:time  + traj +  
                  traj:time  + (time | id),
                  formula.submod = biomk ~ gender + baseline_age + time +
                      (time | id),
                  var.shared = "traj")

summary(res2)

```


### Simpler Model: Drop the random slope in the primary model

Here we drop the random slope in the primary model, which means the `time` 
variable in the primary data set is now a time-dependent variable.

```{r}
res2 <- jlong(data.primod = example_dat$primary,
                  data.submod = example_dat$biomk,
                  formula.primod = score ~ gender + baseline_age + time +
                      baseline_edu + traj + gene + (1 | id),
                  formula.submod = biomk ~ gender + baseline_age + time +
                      (time | id),
                  var.shared = "traj")

summary(res2)
```


[github-url]: https://github.com/ytalau/jlong
